The Host Galaxies of Rapidly Evolving Transients in the Dark Energy Survey

### 2020-04-09 RET_Hosts_SNWG.tex

%define some nice astro shortcuts
\newcommand{\halpha}[0]{H$\alpha$}
\newcommand{\hbeta}[0]{H$\beta$}
\newcommand{\hgamma}[0]{H$\gamma$}
\newcommand{\hdelta}[0]{H$\delta$}
\newcommand{\HII}[0]{$[\textnormal{\textsc{Hii}}]$}
\newcommand{\OII}[0]{$[\textnormal{\textsc{Oii}}]$}
\newcommand{\OI}[0]{$[\textnormal{\textsc{Oi}}]$}
\newcommand{\OIII}[0]{$[\textnormal{\textsc{Oiii}}]$}
\newcommand{\SII}[0]{$[\textnormal{\textsc{Sii}}]$}
\newcommand{\SIII}[0]{$[\textnormal{\textsc{Siii}}]$}
\newcommand{\NII}[0]{$[\textnormal{\textsc{Nii}}]$}
\newcommand{\msun}[0]{M_{\odot}}

\defcitealias{Pursiainen2018}{P18}
\defcitealias{Pursiainen2020}{P20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%This if a fix from https://tex.stackexchange.com/questions/249579/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink-error-with/249743#249743 to sort hyperref breaking in pdfendlink when a reference spills over a page
%\usepackage{etoolbox}
%\makeatletter
%\patchcmd\@combinedblfloats{\box\@outputbox}{\unvbox\@outputbox}{}{%
%  \errmessage{\noexpand\@combinedblfloats could not be patched}%
%}%
%\makeatother
%
%\defcitealias{Smith2019}{S19}
%\defcitealias{Kelly2012}{KK12}
%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

@@

% Abstract of the paper
\begin{abstract}
Rapidly evolving transients (RETs), also termed fast blue optical transients (FBOTs), are a distinct class of astrophysical event. They are characterised by lightcurves that decline much faster than standard supernovae (SNe), span vast ranges in peak luminosity and can be seen to redshifts greater than 1. Their evolution on fast timescales has hindered high quality follow-up observations, such that their origin and explosion/emission mechanism remains unexplained. In this paper, we investigate the host galaxies of the largest RET sample to date from the Dark Energy Survey (DES). Using deep-stacked photometry and emission-lines from OzDES spectroscopy, we derive host galaxy stellar masses and star-formation rates (SFRs) for 49 hosts, and metallicities for 44. We find that RETs explode almost exclusively in star-forming galaxies and are thus likely associated with massive stars. Comparing RET hosts to samples of host galaxies of other explosive transient as well as field galaxies, we find that RETs prefer galaxies with high specific SFRs, indicating a link to young stellar populations, similar to stripped-envelope SNe. RET hosts appear to show a lack of chemical enrichment, their metallicities akin to long duration gamma-ray bursts and superluminous SN host galaxies. There are no clear relationships between properties of the host galaxies and the peak magnitudes or decline rates of the transients themselves.

@@

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
keyword1 -- keyword2 -- keyword3
\end{keywords}

@@

In the standard paradigm of stellar evolution, stars with a zero-age main sequence (ZAMS) mass above $8M_{\sun}$ are believed to explode as a result of a catastrophic collapse of their iron cores and are known as core-collapse supernovae (CCSNe). CCSNe can be split into observationally-determinded subclasses based on their lightcurve and spectral evolution: SNe II display hydrogen features in their spectra \replylluis{that disappear after a few weeks}, and are thought to occur in stars that retain a large fraction of their hydrogen envelope. Conversely, SNe Ib and Ic do not show signatures of hydrogen and are thus referred to collectively as stripped-envelope SNe (SESNe) \replylluis{, having undergone a partial removal of their outer atmospheres}. The SN IIb subclass, which shows hydrogen only at early epochs, is also commonly grouped along with SESNe. \replylluis{SNe IIn, with the n standing for \textit{narrow}, display hydrogen lines with much lower velocities than standard SNe II, indicating their origin from the ejecta impacting on slow-moving circumstellar material (CSM)}. Since the turn of the century, observations of CCSNe, whose lightcurves are primarily powered by the radioactive decay of freshly synthesised Ni-56, have been supplemented by rarer, more exotic transient classes.

@@

 Along with observations of the transients themselves, host galaxies are frequently-used laboratories from which strong inferences about the progenitor stars and explosion mechanisms can be made. CCSNe are confined almost exclusively to galaxies hosting recent or ongoing star formation, due to their origin from massive stars. There are correlations between the expected progenitor mass of different sub-classes of CCSNe and host galaxy properties. On average, SESNe reside in galaxies with higher specific star-formation rate\replychris{s} (sSFRs; \citealt{James2006,Kelly2008}), \replylluis{while studies of the local environments tend to show that SESNe explode closer to \HII~regions than SNe II, indicating that the progenitors are younger and more massive than the various sub-classes of hydrogen-rich SNe II  \citep[e.g.][]{Galbany2018}}. More extreme events tend to occur in galaxies low in mass and high in sSFR, with both GRBs \citep[e.g.][]{Fruchter2006,LeFloch2006,Levesque2010,Kruehler2015,Vergani2015,Perley2016b,Palmerio2019,Taggart2019} and to an even greater degree SLSNe \citep[e.g.][]{Neill2011,Lunnan2014,Leloudas2015,Angus2016,Schulze2018,Taggart2019} exhibiting this association.

@@

Recently, inspection of high-cadence, large-area survey data sets have revealed yet more exotic transients that are less easy to explain with conventional models. \citet{Drout2014} discovered a class of rapidly evolving transients (RETs; also termed `Fast Blue Optical Transients' - FBOTs or `Fast Evolving Luminous Transients' - FELTs) in the Pan-STARRS survey (PS1). \replychris{\citet{Arcavi2016} discovered three rapidly rising, highly luminous objects in the Supernova Legacy Survey (SNLS) and one in the Palomar Transient Factory (PTF), of which one (SNLS04D4ec) declines rapidly like the PS1 sample.} \citet{Pursiainen2018} expanded the known number of RETs to beyond 80 with their sample from the Dark Energy Survey (DES), spanning a redshift range of $\sim 0$ to $>1$. A further sample of five objects has been discovered by the Hyper Suprime-Cam Subaru Strategic Program (SSP) Transient Survey \citep{Tampo2020}. RETs typically rise to peak brightness in less than 10 days, and decline to 10\% of their peak brightness within 30 days, much faster than typical SNe. The photometric measurements of the PS1 and DES RETs seem to be well described by simple expanding blackbodies, although a handful show declining photospheric radii from the outset. Due to the rapid nature of their lightcurves and location at high-redshift, spectral coverage is sparse and signal-to-noise ratio (SNR) is low, such that there has not yet been a conclusive detection of absorption or emission features from the transients and thus the physical mechanism responsible for their rapid evolution remains unexplained.

There are a limited number of events detected in the local Universe whose properties are consistent with the RETs seen in the samples of PS1 and DES at cosmological distances, the most widely studied of which is AT2018cow \citep[e.g.][]{Prentice2018,Perley2019}. The transient declined from its discovery, with constraints on the rise time of 1 day, and across the full range of observed wavelengths did not resemble any known SN, GRB afterglow, or kilonova \replylluis{(KN; \citealt{Ho2019})}. There are many diverse explanations for the power source of AT2018cow touted in the literature, including: magnetars \citep{Mohan2020}; electron capture collapse of merged white dwarfs \citep{Lyutikov2019}; a tidal disruption event (TDE) of a white dwarf \citep{Kuin2019} or of a main sequence star by an intermediate mass black hole \citep{Perley2019}; common envelope jets supernova \replylluis{(CEJSN; \citealt{Soker2019})}; \replychris{or a wind-driven transient \citep{Uno2020}}.
Other nearby rapid transients include the local fast-declining SN-like transient SN2018kzr \citep{McBrien2019} which is explained by the destruction of a white dwarf, and KSN-2015K \citep{Rest2018} whose fast rise and decline is explained by the shock of an SN running into previously-expelled material.  It is currently unclear whether these transients do indeed represent the local analogues of the DES and PS RETs.

In this paper, we present the first comprehensive study of the host galaxies of RETs. We make use of the final DES sample, which builds on \citet{Pursiainen2018} using the final year of data as well as more refined discovery techniques. Using the deep DES photometry from \citet{Wiseman2020} and spectra from OzDES \citep{Lidman2020} we derive host galaxy properties in order to compare them to samples of CCSNe, LGRBs, and SLSNe, as well as the individual local rapid transients. For clarity, we will use the term RET to refer only to events in the high-redshift samples of DES and PS1.

The order of the paper is as follows: \replychris{in Section \ref{sec:sample} we introduce the full DES RET sample and describe the host galaxy observations in Section \ref{sec:obs}. The analysis methods and results are described in Sections \ref{sec:measure} and \ref{sec:analysis} respectively, before a discussion (Section \ref{sec:disc}) and conclusion (Section \ref{sec:conc}).}
Where applicable, we adopt a spatially flat $\Lambda$CDM cosmology with the parameters H$0=70$ km~s$^{-1}$ and $\Omega_{\textrm{M}}=0.3$. Magnitudes are presented in the AB system \citep{Oke1983}.

\begin{figure}
\includegraphics[width=0.5\textwidth]{figs/spec_gr_trise_r_GP5_new_colorscheme.pdf}
\caption{Observer-frame $g-r$ colour \replychris{at maximum light} and \replychris{observer-frame} $r$-band rise-time, derived from Gaussian Processed fits to DES-SN photometry for spectroscopically confirmed SNe and \citetalias{Pursiainen2018} RETs. The location of the red box is designed to maximise the completeness and purity of RETs. \replychris{ The fractions in the legend refer to the number of each class of transient passing the cuts of Section \ref{subsec:new_method} compared to the total number of that class in DES-SN, while the number in parentheses refers to the number inside the red box defining RET parameter space. }
\label{fig:selection}}
\end{figure}

\begin{table*}
\caption{Host galaxy information for the 49 RETs with a host galaxy redshift. A full-length, machine-readable version of this table is available in the online version of the manuscript.}

@@

@@

DES13X3gms     & 35.80095 &  -4.49384 & $23.02 \pm 0.06$ & 0.64792 &      OzDES &       6.6 \\
DES13C1tgd     & 54.06436 & -27.63867 & $20.30 \pm 0.05$ & 0.19647 &      OzDES &      10.2 \\
DES13S2wxf     & 41.61268 &  -0.02634 & $21.31 \pm 0.03$ & 0.56985 &      OzDES &       5.4 \\
DES13X1hav     & 35.03245 &  -5.11022 & $23.63 \pm 0.07$ & 0.58236 &      OzDES &       4.0 \\

@@

DES13X3nyg     & 36.99228 &  -3.91327 & $23.38 \pm 0.06$ & 0.71205 &      OzDES &      45.5 \\
DES13X3gmd     & 36.50409 &  -4.21949 & $22.90 \pm 0.05$ & 0.78082 &      OzDES &      13.9 \\

@@

@@

@@

DES13X2wvv     & 34.86526 &  -6.71603 & $21.89 \pm 0.02$ & 0.47503 &      OzDES &       4.5 \\

@@

DES14X1bnh     & 33.74902 &  -4.79254 & $21.95 \pm 0.03$ & 0.82982 &      OzDES &       3.5 \\
DES14E1aqi     &  8.73970 & -43.40182 & $25.61 \pm 0.23$ &    $^d$ &          - &             - \\
DES15S1fli     & 43.18790 &  -0.88620 & $20.98 \pm 0.02$ & 0.44739 &      OzDES &       2.3 \\
DES14C3gzj     & 52.57168 & -28.14019 & $26.04 \pm 0.10$ &    $^d$ &          - &             - \\
DES14X3pkl     & 37.21103 &  -4.80741 & $22.40 \pm 0.06$ & 0.29537 &      OzDES &       5.2 \\
DES13X3npb     & 36.64218 &  -4.13381 & $20.79 \pm 0.05$ & 0.49542 &      OzDES &       5.2 \\
DES15X2ead     & 36.48913 &  -6.45118 & $20.08 \pm 0.03$ & 0.23175 &      OzDES &       2.0 \\
DES14S2plb     & 41.85667 &  -1.61811 & $18.38 \pm 0.03$ & 0.11531 &      OzDES &       2.0 \\
DES14S2pli     & 41.22837 &  -1.09830 & $20.86 \pm 0.03$ & 0.35478 &      OzDES &       3.4 \\

@@

DES14C3tnz     & 52.86248 & -28.51318 & $22.26 \pm 0.03$ & 0.70452 &      OzDES &       5.8 \\
DES15X3mxf     & 36.74009 &  -5.23991 & $22.43 \pm 0.04$ & 0.43673 &      OzDES &       9.8 \\
DES15X3nlv     & 36.04464 &  -5.02735 & $25.26 \pm 0.08$ &    $^d$ &          - &             - \\
DES15C3mem     & 52.14216 & -29.05851 & $20.28 \pm 0.04$ & 0.61618 &     PRIMUS &               \\
DES15C3lpq     & 52.71204 & -28.61319 & $23.28 \pm 0.03$ & 0.61365 &      OzDES &       6.2 \\
DES15C3nat     & 52.88528 & -28.72370 & $23.39 \pm 0.03$ & 0.83929 &      OzDES &      18.6 \\
DES15C3mgq     & 52.76901 & -28.20882 & $22.97 \pm 0.03$ & 0.23031 &      OzDES &       8.1 \\
DES15E2nqh     &  9.73174 & -43.08707 & $23.29 \pm 0.03$ & 0.51525 &      OzDES &       6.8 \\
DES15C3opk     & 51.66147 & -28.34737 & $23.05 \pm 0.03$ & 0.56984 &      OzDES &      12.2 \\

\bottomrule
\end{tabular}
\begin{tablenotes}
\item[a] Apparent $r$-band Kron magnitude according to DES-SN deep coadds of \citet{Wiseman2020}, not corrected for Galactic foreground reddening.
\item[b] Exposure time only given for spectra which we have used for line measurements rather than just redshift.
\item[c] Host targetted by OzDES but no redshift measurement possible.
\item[d] Host not targetted by OzDES
\item[e] Found by updated search method (Section \ref{subsec:new_method})
\item[f] Found in Season 5 using \citetalias{Pursiainen2018} method 
\end{tablenotes}
\end{threeparttable}
\label{tab:obs}
\end{table*}

\begin{figure*}
\includegraphics[width=0.7\textwidth]{figs/RET_Mosaic.png}
\caption{Selection of DES RET host galaxies in an RGB composite of the DES $gri$ band deep coadds from \citet{Wiseman2020}. The locations of the transients are indicated with cyan crosses. The stamps have a size of 10\arcsec in each direction. \replylluis{DES17X1hjk is considered hostless.}
\label{fig:mosaic}}
\end{figure*}

@@

\section{Sample selection}
\label{sec:sample}
We derive our sample from the 106 RETs discovered in the 5-year DES-SN transient survey. This number expands upon the 72 of \citetalias{Pursiainen2018}. The first reason for the increased sample size is the use of the 5th year of DES-SN, where \citetalias{Pursiainen2018} were only able to make use of the first four years. \replylluis{By imposing the \citetalias{Pursiainen2018} selection criteria on season 5, the sample is increased to 92 objects.} The second reason is an update to the sample selection technique, \replylluis{outlined in the following subsection, which adds a further 14 transients.}

@@

Of objects passing these cuts, the RETs cluster at shorter timescales and bluer colours than other SNe, even in the observer frame. However, using photometric data points directly has several problems that can be improved. For one, measuring peak colour is problematic: DES-SN did not always observe $g$- and $r$-bands on the same or even consecutive nights, thus making it impossible to measure the peak colour in a number of cases. Measuring rise times of 10-15 days is difficult to perform with a one week cadence when it has to be done without fitting a light curve model, hence the rise time values are spread over a wide range. To negate this issue, we use lightcurves interpolated using Gaussian  (GP) as presented in \citet{Pursiainen2019}. The interpolated light curves have a 0.5 day cadence and every epoch \replychris{and band} has a flux value, and an associated uncertainty. Using this technique, SNe Ia and RETs populate two distinct regions of $g-r$ \replylluis{vs.} $t_{\mathrm{rise; r}}$ parameter space (Fig. \ref{fig:selection}), \replychris{where $t_{\mathrm{rise; r}}$ is the time to rise from non-detection to peak $r$-band brightness, and the $g-r$ colour is measured at peak brightness.} We define a region in this parameter space which minimises the contamination of non-RETs (purity) while maximising the total fraction of RETs (completeness). The resulting limits are $-1.5 < g-r < 0.6 $ and $t_{\mathrm{rise; r}} <16.5$, and the parameter space can be seen in Fig. \ref{fig:selection}.

>>>>

%define some nice astro shortcuts
\newcommand{\halpha}[0]{H$\alpha$}
\newcommand{\hbeta}[0]{H$\beta$}
\newcommand{\hgamma}[0]{H$\gamma$}
\newcommand{\hdelta}[0]{H$\delta$}
\newcommand{\hepsilon}[0]{H$\epsilon$}
\newcommand{\HII}[0]{$[\textnormal{\textsc{Hii}}]$}
\newcommand{\OII}[0]{$[\textnormal{\textsc{Oii}}]$}
\newcommand{\OI}[0]{$[\textnormal{\textsc{Oi}}]$}
\newcommand{\OIII}[0]{$[\textnormal{\textsc{Oiii}}]$}
\newcommand{\SII}[0]{$[\textnormal{\textsc{Sii}}]$}
\newcommand{\SIII}[0]{$[\textnormal{\textsc{Siii}}]$}
\newcommand{\NII}[0]{$[\textnormal{\textsc{Nii}}]$}
\newcommand{\msun}[0]{M_{\odot}}

\defcitealias{Pursiainen2018}{P18}
\defcitealias{Wiseman2020}{W20}
\defcitealias{Pursiainen2020}{P20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%This if a fix from https://tex.stackexchange.com/questions/249579/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink-error-with/249743#249743 to sort hyperref breaking in pdfendlink when a reference spills over a page
%\usepackage{etoolbox}
%\makeatletter
%\patchcmd\@combinedblfloats{\box\@outputbox}{\unvbox\@outputbox}{}{%
%  \errmessage{\noexpand\@combinedblfloats could not be patched}%
%}%
%\makeatother
%
%\defcitealias{Smith2019}{S19}
%\defcitealias{Kelly2012}{KK12}
%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

@@

% Abstract of the paper
\begin{abstract}
Rapidly evolving transients (RETs), also termed fast blue optical transients (FBOTs), are a distinct class of astrophysical event. They are characterised by lightcurves that decline much faster than standard supernovae (SNe), span vast ranges in peak luminosity and can be seen to redshifts greater than 1. Their evolution on fast timescales has hindered high quality follow-up observations, such that their origin and explosion/emission mechanism remains unexplained. In this paper we define the largest sample of RETs to date, comprising 106 objects from the Dark Energy Survey. Using deep-stacked photometry and emission-lines from OzDES spectroscopy, we derive host galaxy stellar masses and star-formation rates (SFRs) for 49 host galaxies, and metallicities for 44. We find that RETs explode exclusively in star-forming galaxies and are thus likely associated with massive stars. Comparing RET hosts to samples of host galaxies of other explosive transients as well as field galaxies, we find that RETs prefer galaxies with high specific SFRs, indicating a link to young stellar populations, similar to stripped-envelope SNe. RET hosts appear to show a lack of chemical enrichment, their metallicities akin to long duration gamma-ray bursts and superluminous SN host galaxies. There are no clear relationships between properties of the host galaxies and the peak magnitudes or decline rates of the transients themselves.

@@

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
transients: supernovae -- galaxies: star formation -- galaxies: abundances -- galaxies: photometry 
\end{keywords}

@@

In the standard paradigm of stellar evolution, stars with a zero-age main sequence (ZAMS) mass above $8M_{\sun}$ are believed to explode as a result of a catastrophic collapse of their iron cores and are known as core-collapse supernovae (CCSNe). CCSNe can be split into observationally-determinded subclasses based on their lightcurve and spectral evolution: SNe II display hydrogen features in their spectra \citep{Minkowski1941}, and are thought to occur in stars that retain a large fraction of their hydrogen envelope. Conversely, SNe Ib and Ic do not show signatures of hydrogen \citep[e.g.[]{Filippenko1997} and are thus referred to collectively as stripped-envelope SNe (SESNe), having undergone a partial removal of their outer atmospheres. The SN IIb subclass, which shows hydrogen only at early epochs that disappears after a few weeks \citep{Filippenko1988}, is also commonly grouped along with SESNe. SNe IIn, with the n standing for \textit{narrow}, which refers to the apparent width of the hydrogen emission lines when compared to standard SNe II \citep{Schlegel1990}, indicating their origin from the ejecta impacting on slow-moving circumstellar material (CSM). Since the turn of the century, observations of CCSNe, whose lightcurves are primarily powered by the radioactive decay of freshly synthesised Ni-56, have been supplemented by rarer, more exotic transient classes.

@@

 Along with observations of the transients themselves, host galaxies are frequently-used laboratories from which strong inferences about the progenitor stars and explosion mechanisms can be made. CCSNe are confined almost exclusively to galaxies hosting recent or ongoing star formation, due to their origin from massive stars. There are correlations between the expected progenitor mass of different sub-classes of CCSNe and host galaxy properties. On average, SESNe reside in galaxies with higher specific star-formation rates (sSFRs; \citealt{James2006,Kelly2008}), while studies of the local environments tend to show that SESNe explode closer to \HII~regions than SNe II, indicating that the progenitors are younger and more massive than the various sub-classes of hydrogen-rich SNe II  \citep[e.g.][]{Anderson2012,Galbany2018}. More extreme events tend to occur in galaxies low in mass and high in sSFR, with both GRBs \citep[e.g.][]{Fruchter2006,LeFloch2006,Levesque2010,Kruehler2015,Vergani2015,Perley2016b,Palmerio2019,Taggart2019} and to an even greater degree SLSNe \citep[e.g.][]{Neill2011,Lunnan2014,Leloudas2015,Angus2016,Schulze2018,Taggart2019} exhibiting this association.

@@

Recently, inspection of high-cadence, large-area survey data sets have revealed yet more exotic transients that are less easy to explain with conventional models. \citet{Drout2014} presented a sample of rapidly evolving transients (RETs; also termed `Fast Blue Optical Transients' - FBOTs or `Fast Evolving Luminous Transients' - FELTs) in the Pan-STARRS survey (PS1). \citet{Arcavi2016} reported three rapidly rising, highly luminous objects in the Supernova Legacy Survey (SNLS) and one in the Palomar Transient Factory (PTF), of which one (SNLS04D4ec) declines rapidly like the PS1 sample. \citet[hereafter P18]{Pursiainen2018} expanded the known number of RETs to beyond 80 with their sample from the Dark Energy Survey (DES), spanning a redshift range of $\sim 0$ to $>1$. A further sample of five objects has been discovered by the Hyper Suprime-Cam Subaru Strategic Program (SSP) Transient Survey \citep{Tampo2020}. RETs typically rise to peak brightness in less than 10 days, and decline to 10\% of their peak brightness within 30 days, much faster than typical SNe. The photometric measurements of the PS1 and DES RETs seem to be well described by simple expanding blackbodies, although a handful show declining photospheric radii from the outset. Due to the rapid nature of their lightcurves and location at high-redshift, spectral coverage is sparse and signal-to-noise ratio (SNR) is low, such that there has not yet been a conclusive detection of absorption or emission features from the transients and thus the physical mechanism responsible for their rapid evolution remains unexplained.

There are a limited number of events detected in the local Universe whose properties are consistent with the RETs seen in the samples of PS1 and DES at cosmological distances, the most widely studied of which is AT2018cow \citep[e.g.][]{Prentice2018,Margutti2019,Perley2019}. The transient declined from its discovery, with constraints on the rise time of 1 day, and across the full range of observed wavelengths did not resemble any known SN, GRB afterglow, or kilonova (KN; \citealt{Ho2019}). There are many diverse explanations for the power source of AT2018cow touted in the literature, including: magnetars \citep{Mohan2020}; electron capture collapse of merged white dwarfs \citep{Lyutikov2019}; a tidal disruption event (TDE) of a white dwarf \citep{Kuin2019} or of a main sequence star by an intermediate mass black hole \citep{Perley2019}; common envelope jets supernova (CEJSN; \citealt{Soker2019}); or a wind-driven transient \citep{Uno2020}.
Other nearby rapid transients include the local fast-declining SN-like transient SN2018kzr \citep{McBrien2019} which is explained by the destruction of a white dwarf, and KSN-2015K \citep{Rest2018} whose fast rise and decline is explained by the shock of an SN running into previously-expelled material.  It is currently unclear whether these transients do indeed represent the local analogues of the DES and PS RETs.

In this paper, we present the first comprehensive study of the host galaxies of RETs. We make use of the final DES sample, which builds on \citetalias{Pursiainen2018} using the final year of data as well as more refined discovery techniques. Using the deep DES photometry from \citet[hereafter W20]{Wiseman2020} and spectra from OzDES \citep{Lidman2020} we derive host galaxy properties in order to compare them to samples of CCSNe, LGRBs, and SLSNe, as well as the individual local rapid transients. For clarity, we will use the term RET to refer only to events in the high-redshift samples of DES and PS1.

The order of the paper is as follows: in Section \ref{sec:sample} we introduce the full DES RET sample and describe the host galaxy observations in Section \ref{sec:obs}. The analysis methods and results are described in Sections \ref{sec:measure} and \ref{sec:analysis} respectively, before a discussion (Section \ref{sec:disc}) and conclusion (Section \ref{sec:conc}).
Where applicable, we adopt a spatially flat $\Lambda$CDM cosmology with the parameters $H_0=70$ km~s$^{-1}$ Mpc$^{-1}$ and $\Omega_{\textrm{M}}=0.3$. Magnitudes are presented in the AB system \citep{Oke1983}, and values (uncertainties) are quoted at the 50\textsuperscript{th} (16\textsuperscript{th} and 84\textsuperscript{th}, i.e. $1\sigma$) percentiles of their probability density function.

\begin{figure}
\includegraphics[width=0.5\textwidth]{figs/spec_gr_trise_r_GP5_new_colorscheme.pdf}
\caption{Observer-frame $g-r$ colour at maximum light and observer-frame $r$-band rise-time, derived from Gaussian Processed fits to DES-SN photometry for spectroscopically confirmed SNe and \citetalias{Pursiainen2018} RETs. The location of the red box is designed to maximise the completeness and purity of RETs. The fractions in the legend refer to the number of each class of transient passing the cuts of Section \ref{subsec:new_method} compared to the total number of that class in DES-SN, while the number in parentheses refers to the number inside the red box defining RET parameter space. 
\label{fig:selection}}
\end{figure}

\begin{table*}
\caption{Host galaxy information for the 106 RETs in the DES 5-year sample. A full-length, machine-readable version of this table is available in the online version of the manuscript.}

@@

DES13C1tgd     & 54.06436 & -27.63867 & $20.30 \pm 0.05$ & 0.19647 &      OzDES &      10.17 \\

@@

@@

DES13E2lpk     & 10.09911 & -43.53903 & $20.49 \pm 0.03$ & 0.47541 &      OzDES &       6.00 \\
DES13S2wxf\tnote{e}     & 41.61268 &  $-0.02 \pm 21.3$9 & 0.03266 & 0.56985 &      OzDES &       5.42 \\
DES13X1hav     & 35.03245 &  -5.11022 & $23.63 \pm 0.07$ & 0.58236 &      OzDES &       4.00 \\
DES13X2wvv\tnote{e}     & 34.86526 &  -6.71603 & $21.89 \pm 0.02$ & 0.47503 &      OzDES &       4.50 \\

@@

DES13X3gmd\tnote{e}     & 36.50409 &  -4.21949 & $22.90 \pm 0.05$ & 0.78082 &      OzDES &      13.92 \\
DES13X3gms     & 35.80095 &  -4.49384 & $23.02 \pm 0.06$ & 0.64792 &      OzDES &       6.58 \\

@@

DES13X3npb     & 36.64218 &  -4.13381 & $20.79 \pm 0.05$ & 0.49542 &      OzDES &       5.17 \\
DES13X3nyg     & 36.99228 &  -3.91327 & $23.38 \pm 0.06$ & 0.71205 &      OzDES &      45.50 \\

@@

DES14C3gzj\tnote{e}     & 52.57168 & -28.14019 & $26.04 \pm 0.10$ &    $^d$ &          - &             - \\
DES14C3tnz\tnote{e}     & 52.86248 & -28.51318 & $22.26 \pm 0.03$ & 0.70452 &      OzDES &       5.79 \\
DES14E1aqi\tnote{e}     &  8.73970 & -43.40182 & $25.61 \pm 0.23$ &    $^d$ &          - &             - \\
DES14S2plb     & 41.85667 &  -1.61811 & $18.38 \pm 0.03$ & 0.11531 &      OzDES &       2.00 \\
DES14S2pli     & 41.22837 &  -1.09830 & $20.86 \pm 0.03$ & 0.35478 &      OzDES &       3.42 \\
DES14X1bnh     & 33.74902 &  -4.79254 & $21.95 \pm 0.03$ & 0.82982 &      OzDES &       3.50 \\
DES14X3pkl     & 37.21103 &  -4.80741 & $22.40 \pm 0.06$ & 0.29537 &      OzDES &       5.17 \\
DES15C2eal     & 54.06180 & -29.23000 & $22.99 \pm 0.03$ & 0.22347 &      OzDES &      16.91 \\
DES15C3lpq     & 52.71204 & -28.61319 & $23.28 \pm 0.03$ & 0.61365 &      OzDES &       6.17 \\
DES15C3mem\tnote{e}     & 52.14216 & -29.05851 & $20.28 \pm 0.04$ & 0.61618 &     PRIMUS &               \\
DES15C3mgq     & 52.76901 & -28.20882 & $22.97 \pm 0.03$ & 0.23031 &      OzDES &       8.17 \\
DES15C3nat     & 52.88528 & -28.72370 & $23.39 \pm 0.03$ & 0.83929 &      OzDES &      18.62 \\
DES15C3opk     & 51.66147 & -28.34737 & $23.05 \pm 0.03$ & 0.56984 &      OzDES &      12.17 \\
DES15C3opp     & 51.73962 & -28.11496 & $23.25 \pm 0.04$ & 0.44242 &      OzDES &      16.75 \\
DES15C3pbi     & 52.23620 & -28.00223 & $25.10 \pm 0.05$ &    $^d$ &          - &             - \\
DES15E2nqh     &  9.73174 & -43.08707 & $23.29 \pm 0.03$ & 0.51525 &      OzDES &       6.83 \\
DES15S1fli     & 43.18790 &  -0.88620 & $20.98 \pm 0.02$ & 0.44739 &      OzDES &       2.33 \\
DES15X2ead     & 36.48913 &  -6.45118 & $20.08 \pm 0.03$ & 0.23175 &      OzDES &       2.00 \\

@@

@@

\bottomrule
\end{tabular}
\begin{tablenotes}
\item[a] Apparent $r$-band Kron magnitude according to DES-SN deep coadds of \citetalias{Wiseman2020}, not corrected for Galactic foreground reddening.
\item[b] Exposure time only given for spectra which we have used for line measurements rather than just redshift.
\item[c] Host targetted by OzDES but no redshift measurement possible.
\item[d] Host not targetted by OzDES.
\item[e] Found by updated search method (Section \ref{subsec:new_method}).
\item[f] Found in Season 5 using \citetalias{Pursiainen2018} method.
\item[g] Redshift updated since to \citetalias{Pursiainen2018}.
\end{tablenotes}
\end{threeparttable}
\label{tab:obs}
\end{table*}

\begin{figure*}
\includegraphics[width=0.7\textwidth]{figs/RET_Mosaic.png}
\caption{Selection of DES RET host galaxies in an RGB composite of the DES $gri$ band deep coadds from \citetalias{Wiseman2020}. The locations of the transients are indicated with cyan crosses. The stamps have a size of 10\arcsec in each direction. DES14E2bfx and DES17X1hjk are considered hostless.
\label{fig:mosaic}}
\end{figure*}

@@

\section{Sample selection}
\label{sec:sample}
We derive our sample from the 106 RETs discovered in the 5-year DES-SN transient survey. This number expands upon the 72 of \citetalias{Pursiainen2018}. The first reason for the increased sample size is the use of the 5th year of DES-SN, where \citetalias{Pursiainen2018} were only able to make use of the first four years. By imposing the \citetalias{Pursiainen2018} selection criteria on season 5, the sample is increased to 92 objects. The second reason is an update to the sample selection technique, outlined in the following subsection, which adds a further 14 transients.

@@

Of objects passing these cuts, the RETs cluster at shorter timescales and bluer colours than other SNe, even in the observer frame. However, using photometric data points directly has several problems that can be improved. For one, measuring peak colour is problematic: DES-SN did not always observe $g$- and $r$-bands on the same or even consecutive nights, thus making it impossible to measure the peak colour in a number of cases. Measuring rise times of 10-15 days is difficult to perform with a one week cadence when it has to be done without fitting a light curve model, hence the rise time values are spread over a wide range. To negate this issue, we use lightcurves interpolated using Gaussian  (GP) as presented in \citetalias{Pursiainen2018}. The interpolated light curves have a 0.5 day cadence and every epoch and band has a flux value, and an associated uncertainty. Using this technique, SNe Ia and RETs populate two distinct regions of $g-r$ vs. $t_{\mathrm{rise; r}}$ parameter space (Fig. \ref{fig:selection}), where $t_{\mathrm{rise; r}}$ is the time to rise from non-detection to peak $r$-band brightness, and the $g-r$ colour is measured at peak brightness. We define a region in this parameter space which minimises the contamination of non-RETs (purity) while maximising the total fraction of RETs (completeness). The resulting limits are $-1.5 < g-r < 0.6 $ and $t_{\mathrm{rise; r}} <16.5$, and the parameter space can be seen in Fig. \ref{fig:selection}.

We process all $\sim 30,000$ DES-SN transient candidates with GP. In order to reduce the contamination from active galactic nuclei (AGN), we use a basic convolutional neural network (CNN) classifier. We train the CNN on spectroscopically confirmed SNe of all types and on spectroscopically typed AGN, and use it to separate the sample into two photometric subtypes: AGN-like and SNe-like. The classifier returns SNe-like objects with an accuracy of 0.992 on the test set; the remaining AGN are removed by manual vetting.
The SNe passing the CNN classifier are subjected to LC quality cuts, resulting in 2259 objects, of which 939 lie inside the colour and rise-time region which we defined as RETs. These objects are subject to a further set of cuts. We impose a cut based on a fit of the LC with the \texttt{PSNID} software \citep{Sako2008}. We use thresholds of \texttt{FITPROB}$<0.91$ and \texttt{PBAYES}$<0.82$ to remove highly-probably SNe Ia, which removes 46 objects from the RET parameter space. In order to further remove longer-lived SNe, the decline time to half of the peak brightness must be $<24$ days. This removes 347 SNe, resulting in 546 objects remaining inside the parameter space. The final 546 transients have been visually inspected, with the majority rejected for being spurious detections, obvious multi-season variability that was not picked up by the CNN, or showing evidence for a longer timescale decline.
